---
title: "Replication of the virtual experiment by @Schmitt2019"
output: 
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: bibliography/mee.csl
bibliography: bibliography/library.bib
link-citations: yes
---

Sylvain Schmitt$^1$

*$^1$ CNRS, UMR EcoFoG (Agroparistech, Cirad, INRAE, Université des Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana*

Using the `TROLL` simulator in a long-term virtual experiment, @Schmitt2019 found that functional diversity improves the resilience of tropical forests.
We reproduce a reduced version of this experiment in this vignette to demonstrate how `rcontroll` can help to easily use `TROLL` to implement virtual experiments.
We use the packages `dplyr` & `reshape2` for data processing, `ggplot2` for figures, `FD` for functional divergence indices, and `rcontroll` for simulations.

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
opts_chunk$set(
  echo = T, message = F, warning = F, fig.height = 6, fig.width = 6,
  cache = T, cache.lazy = F)
```

```{r libraries}
suppressMessages(library(dplyr))
# reshape2
library(ggplot2)
suppressMessages(library(FD))
suppressMessages(library(rcontroll))
```

# Simulation experiment

## Experimental design

We simulate $N=10$ communities for each species richness level through $SR\in[5,10,20]$ over 600 years,
resulting in 6 communities with species randomly sampled in `TROLLv3_species`.

```{r design}
N <- 10
SR <- c(5, 10, 20)
data("TROLLv3_species")
communities <- lapply(rep(SR, each = N),
                      function(n) sample_n(TROLLv3_species, n)) 
names(communities) <- paste0("com", 1:length(communities))
```

## Pre-disturbance simulations

Once the communities were defined, we use the `stack` function of `rcontroll` to simulate 6 forests from the bare ground before disturbance.
We use the the `autoplot` function of `rcontroll` to represent spatial pattern of the forest before disturbance in the `com1` community (Fig. \@ref(fig:predistsimfig)).

```{r predistsim, eval=F}
simulations_pre <- stack(name = "pre", 
      simulations = communities_stats$community,
      path = "simulations",
      global = generate_parameters(nbiter = 12*600), 
      species = bind_rows(communities, .id = "simulation"), 
      climate = TROLLv3_climatedaytime12, 
      daily = TROLLv3_daytimevar, 
      verbose = FALSE,
      cores = 30)
save(simulations_pre, file = "simulations/simulations_pre.Rdata")
```

```{r predistsimfig, fig.cap="Spatial pattern of the forest before disturbance in the `com1` community. Trees are represented by points with size corresponding to diameter at breast height (DBH, m) and colour to height (m)."}
load("simulations/simulations_pre.Rdata")
subsims <- simulations_pre
subsims@forest <- filter(subsims@forest, simulation == "com1")
rcontroll::autoplot(subsims, what = "spatial", variable = "height") +
  viridis::scale_color_viridis()
```

## Disturbance simulations

@Schmitt2019 used a particular type of random disturbance as a first test case to explore the response of forest biodiversity on tropical forest resilience to disturbance.
Consequently, we can simply use the `sample_frac` function from `dplyr` to reduce the forest to a random defined fraction per simulation thanks to the `group_by` function from `dplyr`.
This results in the spatial pattern of the forest at disturbance in the `com1` community shown in Fig. \@ref(fig:distsimfig). 
However, future studies could build on this random disturbance to include more mechanistic details and explore the effect of different disturbance types with homemade R code or specific packages (*e.g.* [`LoggingLab`](https://github.com/VincyaneBadouard/LoggingLab)).

```{r distsim}
simulations_pre@forest <- simulations_pre@forest %>% 
  group_by(simulation) %>% 
  sample_frac(3/4)
```

```{r distsimfig, fig.cap="Spatial pattern of the forest at disturbance in the `com1` community. Trees are represented by points with size corresponding to diameter at breast height (DBH, m) and colour to height (m)."}
subsims <- simulations_pre
subsims@forest <- filter(subsims@forest, simulation == "com1")
rcontroll::autoplot(subsims, what = "spatial", variable = "height") +
  viridis::scale_color_viridis()
```

## Post-disturbance simulations

Once the communities were disturbed, we use once again the `stack` function of `rcontroll` to simulate the post-diturbance trajectories of the 6 forests.
We use again the the `autoplot` function of `rcontroll` to represent spatial pattern of the forest after disturbance in the `com1` community (Fig. \@ref(fig:postdistsimfig)).

```{r postdistsim, eval=F}
simulations_post <- stack(name = "post", 
      simulations = communities_stats$community,
      path = "simulations",
      global = simulations_pre@inputs$global, 
      species = bind_rows(communities, .id = "simulation"), 
      climate = TROLLv3_climatedaytime12, 
      daily = TROLLv3_daytimevar, 
      forest = as.data.frame(simulations_pre@forest),
      verbose = FALSE,
      cores = 30)
save(simulations_post, file = "simulations/simulations_post.Rdata")
```

```{r postdistsimfig, fig.cap="Spatial pattern of the forest after disturbance in the `com1` community. Trees are represented by points with size corresponding to diameter at breast height (DBH, m) and colour to height (m)."}
load("simulations/simulations_post.Rdata")
subsims <- simulations_post
subsims@forest <- filter(subsims@forest, simulation == "com1")
rcontroll::autoplot(subsims, what = "spatial", variable = "height") +
  viridis::scale_color_viridis()
```

## Undisturbed simulations

For control, we use once again the `stack` function of `rcontroll` to simulate the undisturbed trajectories of the 6 forests.
We use again the the `autoplot` function of `rcontroll` to represent spatial pattern of the undisturbed forest in the `com1` community (Fig. \@ref(fig:undistsimfig)).

```{r undistsim, eval=F}
load("simulations/simulations_pre.Rdata")
simulations_undist <- stack(name = "undist", 
      simulations = communities_stats$community,
      path = "simulations",
      global = simulations_pre@inputs$global, 
      species = simulations_pre@inputs$species, 
      climate = TROLLv3_climatedaytime12, 
      daily = TROLLv3_daytimevar, 
      forest = as.data.frame(simulations_pre@forest),
      verbose = FALSE,
      cores = 30)
save(simulations_undist, file = "simulations/simulations_undist.Rdata")
```

```{r undistsimfig, fig.cap="Spatial pattern of the undisturbed forest in the `com1` community. Trees are represented by points with size corresponding to diameter at breast height (DBH, m) and colour to height (m)."}
load("simulations/simulations_undist.Rdata")
subsims <- simulations_undist
subsims@forest <- filter(subsims@forest, simulation == "com1")
rcontroll::autoplot(subsims, what = "spatial", variable = "height") +
  viridis::scale_color_viridis()
```

# Analyses

## Quantifying simulated community diversity

Specific and functional diversity are assessed for each simulation just before disturbance,
i.e. at the end of the pre-disturbance simulation stage,
with indices of specific richness ($SR$) and functional diversity 
($FDiv$, the volume of the functional space occupied by the community; $FEve$, the regularity of the distribution of abundance in this volume) 
using the `dbFD` function from `FD`.
This results in the values shown in Fig. \@ref(fig:prediststatsfig).

```{r prediststats, include=FALSE}
communities_stats <- left_join(simulations_pre@inputs$species, 
          filter(simulations_pre@species, iter == max(iter)),
          by = c("simulation" = "simulation", "s_name" = "species")) %>% 
  group_by(simulation, s_name) %>% 
  sample_n(1) %>% 
  group_by(simulation) %>% 
  mutate(sum1 = ifelse(is.na(sum1), 0, sum1)) %>% 
  mutate(abundance = sum1/sum(sum1)) %>% 
  select(simulation, s_name, s_LMA, s_Nmass, s_Pmass, s_wsg, s_dbhmax, s_hmax, s_ah, abundance) %>% 
  filter(abundance > 0) %>% 
  split(.$simulation) %>% 
  lapply(function(tab) {
    tab <- as.data.frame(arrange(tab, s_name))
        print(tab$s_name)
    row.names(tab) <- tab$s_name
    dbFD(x = tab[c("s_LMA", "s_Nmass", "s_Pmass", "s_wsg", "s_dbhmax", "s_hmax", "s_ah")], 
         a = reshape2::dcast(tab, 1 ~ s_name, value.var = "abundance")[-1], m = "max", calc.CWM = F)
  }) %>% 
    lapply(as.data.frame) %>% 
  bind_rows(.id = "community")
unlink("vert.txt") 
```

```{r prediststats2, eval=F}
communities_stats <- left_join(simulations_pre@inputs$species, 
                               filter(simulations_pre@species, iter == max(iter)),
                               by = c("simulation" = "simulation", "s_name" = "species")) %>% 
  group_by(simulation) %>% 
  mutate(sum1 = ifelse(is.na(sum1), 0, sum1)) %>% 
  mutate(abundance = sum1/sum(sum1)) %>% 
  select(simulation, s_name, s_LMA, s_Nmass, s_Pmass, s_wsg, s_dbhmax, s_hmax, s_ah, abundance) %>% 
  filter(abundance > 0) %>% 
  split(.$simulation) %>% 
  lapply(function(tab) {
    tab <- as.data.frame(arrange(tab, s_name))
    print(tab$s_name)
    row.names(tab) <- tab$s_name
    dbFD(x = tab[c("s_LMA", "s_Nmass", "s_Pmass", "s_wsg", "s_dbhmax", "s_hmax", "s_ah")], 
         a = reshape2::dcast(tab, 1 ~ s_name, value.var = "abundance")[-1], m = "max", calc.CWM = F)
  }) %>% 
  lapply(as.data.frame) %>% 
  bind_rows(.id = "community")
```

```{r prediststatsfig, fig.cap="Functional richness (FRic) and evenness (FEve) across simulated communities pre-disturbance with varying species richness (SR, shown by boint shape)."}
ggplot(communities_stats, 
       aes(FRic, FEve, shape = as.factor(nbsp))) +
  geom_point() +
  theme_bw() +
  scale_shape_discrete("Species\nrichness\n(SR)") +
  xlab("Functional richness (FRic)") +
  ylab("Functional evenness (FEve)") 
```

## Ecosystem recovery from disturbance

We characterise changes in ecosystem properties of simulated tropical forests with respect to forest structure 
(above-ground biomass, AGB in $MgC.ha^{-1}$ ; basal area, BA in $m^2.ha^{-1}$ ; total number of stems, $N$ ; 
number of stems over 10 cm dbh, $N10$ ; and number of stems over 30 cm dbh, $N30$).
The recovery of the simulated forest at a time $t$ regarding a given property $X$, $Rx(t)$, 
is defined relatively to a non-disturbed baseline [@Schmitt2019]:

$$R_x(t)=\frac{X_D(t)}{X_C(t)}$$

where $X_D(t)$ and $X_C(t)$ are the values for the ecosystem property $X$ at time t in the disturbed simulation and in the control simulation, respectively. 
To capture the temporal dimension of resilience,  we integrate $d_x(t) = \sqrt{(R_{eq}-R[X(t)])^2}$ through time which quantifies the cumulative difference between disturbed and control runs.
We finally compute the inverse of this area as an index of ecosystem resilience, 
henceforth denoted $I_R$ , so that a higher $I_R$ reflects a higher resilience:

$$I_R(X)=(\int _{t=disturbiter} ^{t=disturbiter+600} d_x(t).dt)^{-1}$$

Distance to equilibrium $d_{eq}$ and resilience index $I_R$ are calculated in a multi-dimensional space for forest structure (AGB, BA, N, N10, and N30).
This results in the values shown in Fig. \@ref(fig:resfig).

```{r red}
resilience <- left_join(
  select(simulations_post@ecosystem, simulation, iter, agb, ba, sum1, sum10, sum30) %>% 
    reshape2::melt(c("simulation", "iter"), value.name = "X_D"),
  select(simulations_undist@ecosystem, simulation, iter, agb, ba, sum1, sum10, sum30) %>% 
    reshape2::melt(c("simulation", "iter"), value.name = "X_C")
) %>% 
  mutate(R = X_D/(X_C+10^-6)) %>% 
  group_by(simulation, iter) %>% 
  summarise(d_x = sqrt(sum((1-R)^2))) %>% 
  group_by(simulation) %>% 
  summarise(I_R = 1/sum(d_x))
```

```{r resfig, fig.cap="Variation of forest resilience across simulations. Resilience index $I_R$ was represented for each simulation."}
ggplot(resilience, aes(simulation, I_R)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  ylab(expression(I[R])) +
  theme(axis.title.y = element_blank())
```

## Biodiversity effect

We test the influence of three factors on the resilience index $I_R$:
pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) as follows:

$$log(I_R) \sim \mathcal N (\theta_{SR}+\theta_{FEve}\times FEve+\theta_{FDiv}\times FDiv;\sigma^2)$$

where $\theta_{SR}$ is a vector of 3 parameters corresponding to the three species richness levels, 
$\theta_{FEve}$ and $\theta_{FDiv}$ respectively model the effect of $FEve$ and $FDiv$ on $I_R$  under a normal law of variance $\sigma^2$. 
The ecosystem resilience index $I_R$ was log-transformed to improve residuals’ normality.
We find a significant effect of $FEve$ and $FDiv$ on $I_R$ as shown in the regression table resulting in the relation shown in Fig. \@ref(fig:fig).

```{r data}
data <- communities_stats %>% 
  left_join(resilience, by = c("community" = "simulation")) %>% 
  mutate(SR = nbsp) %>% 
  select(community, I_R, SR, FDiv, FEve)
```

```{r lm}
lm(log(I_R) ~ SR + FEve + FDiv, data) %>% 
  sjPlot::tab_model()
```

```{r fig, fig.cap="Variation of forest resilience with taxonomic and functional diversity. Resilience index $I_R$ was represented against functional diversity ($FDiv$); dot size varies with species richness ($SR$) whereas dot color indicates functional evenness ($FEve$). Solid line and grey area show fitted relationships with their mean and variance of the form $I_R \\sim \\mathcal N (\\theta_{FDiv}\\times FDiv;\\sigma^2)$."}
ggplot(data, aes(FDiv, I_R)) +
  geom_smooth(method = "lm", col = "grey", fill = "lightgrey") +
  geom_point(aes(col = FEve, size = SR)) +
  theme_bw() +
  viridis::scale_color_viridis("Functional\nevenness\n(FEve)") +
  xlab("Functional diversity (FDiv)") +
  ylab("Resilience indice Ir") +
  scale_size_continuous("Species\nrichness\n(SR)")
```

# Conclusion

Using a reduced virtual experiment, we find back the results from @Schmitt2019: functional diversity improves the resilience of tropical forests.
However, we used far less code and time than in the original study using homemade code with an older version of `TROLL` (S; Schmitt pers. com.).

# References
